---
title: "summarizeTumor Package Documentation"
author: "Poorna Murali"
date: "`r Sys.Date()`"
output:
  github_document:
    toc: true
    toc_depth: 2
    number_sections: true
---



```{r, include = FALSE}
# Set CRAN mirror to avoid errors
options(repos = c(CRAN = "https://cran.rstudio.com"))
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# Introduction

**`summarizeTumor`Package Description**

The `summarizeTumor` package provides a set of functions designed to help users summarize cancer-related tumor data. Specifically, it calculates statistical summaries (e.g., mean, median, standard deviation) for numeric columns within a cancer dataset, grouped by tumor type (diagnosis). This package is particularly useful for gaining insights into the distribution of tumor-related metrics and facilitates data pre-processing and analysis.

**Key Functionality**

The main function, summarize_tumor_data(), allows users to summarize numeric columns (e.g., radius_mean, texture_mean) in the dataset by different tumor types or diagnoses. The function supports flexibility in output, allowing for customized aggregation and data grouping.

# Installation

To install the summarizeTumor package from GitHub, you’ll need to use the devtools package. Here’s how to do it:

``` {r}
# Install devtools if not already installed
install.packages("devtools")
devtools::install_github("https://github.com/stat545ubc-2024/summarizeTumor.git")
```
**Demonstrated Usage**

Once installed, you can start using the `summarize_tumor_data()` function as shown below. The function’s parameters include:

`data`: A data frame or tibble containing the tumor data (e.g., the cancer dataset with columns like `radius_mean`, `texture_mean`).
`group_col`: The column to group by when summarizing the data (e.g., `diagnosis`, which indicates whether the tumor is benign or malignant).
`column_to_summarize`: The numeric column to calculate statistics for (e.g., `radius_mean`, `texture_mean`).

# Examples

Each row represents a sample with features such as `radius_mean`, `texture_mean`, `perimeter_mean`, etc., and a column `diagnosis` representing the tumor type (benign or malignant).

**Basic Usage**

Summarize the radius_mean in the cancer_sample dataset, grouped by diagnosis (benign or malignant):

```{r example}
library(summarizeTumor)
library(dplyr)
# Load the cancer_sample dataset
data("cancer_sample", package = "datateachr")

# Summarize the 'radius_mean' column by diagnosis
tumor_summary <- summarize_tumor_data(cancer_sample, "radius_mean")

# Print the summarized data
print(tumor_summary)
```

**Plotting**

You can also visualize the distribution of the radius_mean values for each diagnosis group using a boxplot.

```{r pressure, echo = FALSE}
# Create a boxplot of radius_mean grouped by diagnosis
library(ggplot2)

ggplot(cancer_sample, aes(x = diagnosis, y = radius_mean)) +
  geom_boxplot() +
  labs(title = "Distribution of Radius Mean by Diagnosis", 
       x = "Diagnosis", 
       y = "Radius Mean") +
  theme_minimal()

```

This plot shows the distribution of `radius_mean` values for benign and malignant tumor samples in the dataset. The boxplot allows you to visually compare the central tendency and spread of the radius_mean variable across different diagnosis categories.

# License

This package is licensed under the MIT License.



